---
title: "Cumulative infections over 25 years"
author: "Mathew Kiang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Static tables 

```{r data, results='hide'}
## Imports ----
library(tidyverse)
library(duckdb)
library(here)
library(kableExtra)
library(knitr)
source(here("code", "utils.R"))

## Helper function ----
prettyround <- function(x, accuracy_x = 0.1) {
    scales::label_number(
        accuracy = accuracy_x,
        scale_cut = scales::cut_short_scale()
    )(x)
}

## Data ----
us_final <- pull_summary_data(
    state_x = "US",
    vaccine_x = c(1, .9, 1.05, .5, .75),
    time_max = 365 * 25
) |>
    filter(
        time == 365 * 25 - 1,
        metric == "cume_new_infectious"
    ) |>
    mutate(mean = winsorized_mean) |>
    select(-starts_with("winsorized"))
```


## Cumulative cases across all outcomes (**rounded**)

Cumulative number of cases at 25 years, by pathogen and vaccine coverage. 

```{r}
us_final |>
    rowwise() |>
    transmute(
        pathogen_cat,
        vaccine_coverage_cat,
        estimate_mean95 = sprintf(
            "%s (95%% UI: %s-%s)",
            prettyround(mean),
            prettyround(p025),
            prettyround(p975)
        ),
        estimate_medianiqr = sprintf(
            "%s (IQR: %s-%s)",
            prettyround(median),
            prettyround(p250),
            prettyround(p750)
        )
    ) |>
    arrange(pathogen_cat, desc(vaccine_coverage_cat)) |>
    kable(format = "html") |>
    kable_styling("striped", full_width = TRUE)
```

## Additional cases compared to current levels (**rounded**; 50% vaccine scenario)

How many **more** cases would happen after 25 years, compared to the expected number of cases (under current vaccine coverage)?

```{r}
## Constants ----
TARGET_VACCINE <- .5
us_final |>
    filter(vaccine_coverage %in% c(1, TARGET_VACCINE)) |>
    select(pathogen_cat, vaccine_coverage, mean:p975) |>
    pivot_wider(values_from = mean:p975, names_from = vaccine_coverage) |>
    transmute(
        pathogen_cat,
        additional_cases_mean95 =
            sprintf(
                "%s (95%% UI: %s-%s)",
                prettyround(`mean_0.5` - mean_1),
                prettyround(`p025_0.5` - p025_1),
                prettyround(`p975_0.5` - p975_1)
            ),
        additional_cases_medianiqr =
            sprintf(
                "%s (IQR: %s-%s)",
                prettyround(`median_0.5` - median_1),
                prettyround(`p250_0.5` - p250_1),
                prettyround(`p750_0.5` - p750_1)
            )
    ) |>
    arrange(pathogen_cat) |>
    kable(format = "html") |>
    kable_styling("striped", full_width = TRUE)
```


## Cumulative cases across all outcomes (**exact**)
```{r}
us_final |>
    transmute(
        pathogen_cat,
        vaccine_coverage_cat,
        estimate_mean95 = sprintf(
            "%s (95%% UI: %s-%s)",
            prettyNum(round(mean), big.mark = ","),
            prettyNum(round(p025), big.mark = ","),
            prettyNum(round(p975), big.mark = ",")
        ),
        estimate_medianiqr = sprintf(
            "%s (IQR: %s-%s)",
            prettyNum(round(median), big.mark = ","),
            prettyNum(round(p250), big.mark = ","),
            prettyNum(round(p750), big.mark = ",")
        )
    ) |>
    arrange(pathogen_cat, desc(vaccine_coverage_cat)) |>
    kable(format = "html") |>
    kable_styling("striped", full_width = TRUE)
```

## Additional cases compared to current levels (**exact**; 50% vaccine scenario)

```{r}
## Constants ----
TARGET_VACCINE <- .5
us_final |>
    filter(vaccine_coverage %in% c(1, TARGET_VACCINE)) |>
    select(pathogen_cat, vaccine_coverage, mean:p975) |>
    pivot_wider(values_from = mean:p975, names_from = vaccine_coverage) |>
    transmute(
        pathogen_cat,
        additional_cases_mean95 =
            sprintf(
                "%s (95%% UI: %s-%s)",
                prettyNum(round(`mean_0.5` - mean_1), big.mark = ","),
                prettyNum(round(`p025_0.5` - p025_1), big.mark = ","),
                prettyNum(round(`p975_0.5` - p975_1), big.mark = ",")
            ),
        additional_cases_medianiqr =
            sprintf(
                "%s (IQR: %s-%s)",
                prettyNum(round(`median_0.5` - median_1), big.mark = ","),
                prettyNum(round(`p250_0.5` - p250_1), big.mark = ","),
                prettyNum(round(`p750_0.5` - p750_1), big.mark = ",")
            )
    ) |>
    arrange(pathogen_cat) |>
    kable(format = "html") |>
    kable_styling("striped", full_width = TRUE)
```

# Interactive table, full results

```{r data2, results='hide'}
## Imports ----
library(DT)

## Data ----
us_final <- pull_summary_data(
    state_x = "US",
    time_max = 365 * 30
) |>
    filter(
        time == 365 * 25 - 1,
        metric == "cume_new_infectious"
    ) |>
    mutate(mean = winsorized_mean) |>
    select(-starts_with("winsorized"))
```


```{r}
us_final |>
    rowwise() |>
    transmute(
        pathogen_cat,
        vaccine_coverage_cat,
        estimate_mean95 = sprintf(
            "%s (95%% UI: %s-%s)",
            prettyround(mean),
            prettyround(p025),
            prettyround(p975)
        ),
        estimate_medianiqr = sprintf(
            "%s (IQR: %s-%s)",
            prettyround(median),
            prettyround(p250),
            prettyround(p750)
        )
    ) |>
    arrange(pathogen_cat, desc(vaccine_coverage_cat)) |>
    DT::datatable(
                colnames = c(
                    "Pathogen",
                    "Vaccine scenario",
                    "Mean (95% UI)",
                    "Median (IQR)"
                ),
                rownames = FALSE,
                filter = list(
                    position = 'top'
                ),
                options = list(
                    columnDefs = list(list(targets = c(2, 3), 
                                           searchable = FALSE)),
                    paging = TRUE,
                               searching = TRUE,
                               pageLength = 100,
                searchCols = list(
                    NULL, 
                                  list(search = '["5% higher","Current levels","10% lower","20% lower","50% lower"]'))))
```

