---
title: "Demographic details"
author: "Mathew Kiang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r data, results='hide'}
## Imports ----
library(tidyverse)
library(duckdb)
library(here)
library(kableExtra)
library(knitr)
source(here("code", "utils.R"))

contact_matrix <- utils::read.csv(here::here("data", "prem2021_v2.csv"), header = FALSE)
```

```{r}
## Modified simulation function 
simulate_outbreak <- function(R0,
                              gamma,
                              sigma,
                              lambda_import,
                              initial_immune,
                              birth_rate,
                              age_population,
                              age_specific_mu_rate,
                              contact_matrix_load,
                              pathogen_x = "rubella",
                              target_coverage = 0,
                              vaccine_efficacy,
                              transmission_reduction,
                              static_importation = FALSE,
                              import_by_population,
                              days = 365 * 25) {
    ## Constants
    beta_start <- R0 * gamma
    aging_fraction <- 1 / (5 * 12) # Monthly aging for compartments
    age_bins <- seq(0, 85, 5)
    N_AGES <- NROW(age_bins)
    
    female_df <- dplyr::tribble(
    ~age_group, ~female, ~total,
    15, 30958568, 63317757,
    20, 32319438, 66360962,
    25, 34743499, 71103584,
    30, 33015027, 66966208,
    35, 32289146, 64830919,
    40, 29913869, 59610918
) |>
    dplyr::transmute(age_group,
        prob_female = female / total
    )

    # Vaccination parameters
    # NOTE: For polio and diphtheria, we assume 100% seroconversion
    # so perfect vaccine efficacy at any given coverage, but each have
    # reduced transmission due to vaccination
    effective_coverage <- min(target_coverage, 1) * vaccine_efficacy

    # Exponential movement rate for 5-year compartment with daily time step
    # (only used when effective coverage > 0 and only for first age bin)
    aging_rate <- 1 / (5 * 365)

    ## Age params
    age_distribution <- age_population / sum(age_population)
    age_specific_mu <- 1 - exp(-age_specific_mu_rate)

    # Calculate contact matrix
    contact_matrix <- as.matrix(t(contact_matrix_load))
    # Compute next generation matrix
    G_trial <- beta_start / gamma * t(contact_matrix_load)
    # Calculate dominant eigenvalue of the trial matrix
    lambda_max_trial <- max(Re(eigen(G_trial)$values))
    # Scale beta to match desired R0, to account for mixing matrix
    beta <- beta_start * (R0 / lambda_max_trial)

    # Initialize compartments for each age group
    S <- as.integer((1 - initial_immune) * age_population)
    E <- rep(0, N_AGES)
    I <- rep(0, N_AGES)
    R <- as.integer(initial_immune * age_population)

    # Results storage containers
    susceptible <- matrix(0, nrow = days, ncol = N_AGES)
    exposed <- matrix(0, nrow = days, ncol = N_AGES)
    infectious <- matrix(0, nrow = days, ncol = N_AGES)
    recovered <- matrix(0, nrow = days, ncol = N_AGES)
    new_infectious_store <- matrix(0, nrow = days, ncol = N_AGES)
    vrt_store <- matrix(0, nrow = days, ncol = N_AGES)
    imported_store <- matrix(0, nrow = days, ncol = N_AGES)
    
    ## The storage containers we actually want
    births_store <- vector("numeric", length = days)
    deaths_store <- vector("numeric", length = days)
    population_store <- vector("numeric", length = days)
    sus_repro_female_store <- vector("numeric", length = days)

    # Note that we track vaccination reduced transmission for *all* outcomes
    # but it only changes for polio and diphtheria. For other pathogens, it's
    # always 0. Just makes reassembling the results easier since the output
    # will be consistent.
    VRT_number <- rep(0, N_AGES)
    if (pathogen_x == "polio") {
        # Since IPV roll out in 2000, all seropositive during this
        # period is likely due to IPV vaccination (approx ages 1-25 years)
        VRT_number[1:5] <- R[1:5]
        S[1:5] <- S[1:5] + R[1:5] # Remove IPV protection against infection
        R[1:5] <- R[1:5] - R[1:5] # Should be zero
    }

    if (pathogen_x == "diphtheria") {
        VRT_number <- R
        S <- S + R # Remove protection against infection
        R <- R - R # Should be zero
    }
    
    # Record susceptible fraction at time 0
    susceptible_fraction_t0 <- sum(S) / sum(S + E + I + R)

    # Simulation loop
    for (t in seq.int(days)) {
        # Store results
        susceptible[t, ] <- S
        exposed[t, ] <- E
        infectious[t, ] <- I
        recovered[t, ] <- R
        vrt_store[t, ] <- VRT_number
        
        population_store[t] <- sum(S + E + I + R)
        sus_repro_female_store[t] <- sum((S[4:9] - VRT_number[4:9]) * female_df$prob_female)

        # frequency-dependent transmission (age-based mixing)
        current_pop_size <- S + E + I + R
        if (pathogen_x %in% c("polio", "diphtheria")) {
            VRT_coverage <- VRT_number / current_pop_size
            VRT_reduction_transmission <- transmission_reduction * VRT_coverage
            force_of_infection <- beta * (contact_matrix %*% (I * (1 - VRT_reduction_transmission) / current_pop_size))
        } else {
            force_of_infection <- beta * (contact_matrix %*% (I / current_pop_size))
        }

        # Calculate transition probabilities for each age group (from rates)
        p_SE <- 1 - exp(-force_of_infection) # S -> E
        p_EI <- 1 - exp(-sigma) # E -> I
        p_IR <- 1 - exp(-gamma) # I -> R

        # Account for maternal protection factor (6 months of perfect immunity after birth)
        # Based on 6 months, within a 5 year age bin
        # (10% within this age group at protected by maternal immunity)
        maternal_protection_factor <- 0.1
        # Remove subset from S state to account for maternal protection
        S_maternalimmunity <- S
        S_maternalimmunity[1] <- round(S_maternalimmunity[1] *
            (1 - maternal_protection_factor))

        # Stochastic transitions for each age group
        new_exposed <- stats::rbinom(N_AGES, S, p_SE)
        new_infectious <- stats::rbinom(N_AGES, E, p_EI)
        new_recovered <- stats::rbinom(N_AGES, I, p_IR)

        # Importation process
        if (static_importation) {
            lambda_import_t <- lambda_import
        } else {
            susceptible_fraction_t <- sum(S) / sum(S + E + I + R)
            importation_multiplier <- susceptible_fraction_t / susceptible_fraction_t0
            lambda_import_t <- lambda_import * importation_multiplier
        }
        
        prob_weights <- age_distribution
        if (!import_by_population) {
            prob_weights <- (S / (S + E + I + R)) * age_distribution
        }

        import_cases_dt <- stats::rpois(1, lambda_import_t)
        imported_cases_age <- sample(
            x = c(1:N_AGES),
            size = import_cases_dt,
            prob = prob_weights,
            replace = TRUE
        )
        imported_cases <- tabulate(
            factor(imported_cases_age, levels = 1:N_AGES),
            nbins = N_AGES
        )
        imported_store[t, ] <- imported_cases

        # Update compartments
        S <- S - new_exposed
        E <- E + new_exposed - new_infectious
        I <- I + new_infectious - new_recovered + imported_cases
        R <- R + new_recovered

        # count incident cases (including imports)
        new_infectious_store[t, ] <- new_infectious + imported_cases

        # Apply age-specific death rates
        total_deaths <- stats::rbinom(N_AGES, S + E + I + R, age_specific_mu)
        if (sum(total_deaths) > 0) {
            deaths_S <- round((S / (S + E + I + R)) * total_deaths)
            deaths_E <- round((E / (S + E + I + R)) * total_deaths)
            deaths_I <- round((I / (S + E + I + R)) * total_deaths)
            deaths_R <- total_deaths - (deaths_S + deaths_E + deaths_I)
            # Ensure all deaths are distributed

            S <- S - deaths_S
            E <- E - deaths_E
            I <- I - deaths_I
            R <- R - deaths_R
        }
        
        deaths_store[t] <- sum(total_deaths)

        # Apply births
        if (pathogen_x %in% c("polio", "diphtheria")) {
            births <- stats::rpois(1, sum(S + E + I + R) * birth_rate)
            VRT_number[1] <- VRT_number[1] + round(births * effective_coverage)
            S[1] <- round(S[1] + births)

            ## Make sure VRT number doesn't exceed number of susceptibles
            VRT_number <- pmin(VRT_number, S)
        } else {
            # Births added to the youngest age group
            births <- stats::rpois(1, sum(S + E + I + R) * birth_rate)
            S[1] <- round(S[1] + births * (1 - effective_coverage))
            R[1] <- round(R[1] + births * effective_coverage)
        }
        
        births_store[t] <- births

        # Age population progression
        # (individuals age by shifting compartments every month)
        if (t %% 30 == 0) {
            # Perform aging every month
            S_next <- round(S * aging_fraction)
            E_next <- round(E * aging_fraction)
            I_next <- round(I * aging_fraction)
            R_next <- round(R * aging_fraction)

            # Removed aged individuals
            S <- S - c(S_next[-length(S)], 0)
            E <- E - c(E_next[-length(E)], 0)
            I <- I - c(I_next[-length(I)], 0)
            R <- R - c(R_next[-length(R)], 0)

            # Add aged individuals to the next bin
            S <- S + c(0, S_next[-length(S)])
            E <- E + c(0, I_next[-length(E)])
            I <- I + c(0, E_next[-length(I)])
            R <- R + c(0, R_next[-length(R)])

            # Add aging to the IPV vaccine tracker (number of IPV vaccinated persons per age group)
            if (pathogen_x %in% c("polio", "diphtheria")) {
                # Aging of vaccine coverage
                VRT_next <- round(VRT_number * aging_fraction)
                VRT_number <- VRT_number - c(VRT_next[-length(VRT_number)], 0)
                VRT_number <- VRT_number + c(0, VRT_next[-length(VRT_number)])
            }
        }
    }

    ## Remove this here so cumsum() below calculates the right thing
    rm(new_infectious)

    # Create a data frame for plotting
    results <- data.frame(
        time = 1:days,
        population = population_store,
        births = births_store,
        deaths = deaths_store,
        female_repro_sus = sus_repro_female_store
    ) 

    results
}

```

```{r}
holder <- map_dfr(.x = c("DC", state.abb),
    .f = ~ {
        rubella_params <- return_pathogen_params(.x, pathogen_x = "rubella")

        R0 <- rubella_params$R0
        gamma <- rubella_params$gamma
        sigma <- rubella_params$sigma
        lambda_import <- rubella_params$lambda_import
        initial_immune <- return_initial_immunity(.x, pathogen_x = "rubella")
        birth_rate <- return_birth_rate(.x) / 365
        age_population <- return_age_structure(.x)
        age_specific_mu_rate <- return_death_rate(.x) / 365
        contact_matrix_load <- contact_matrix
        vaccine_efficacy <- rubella_params$vaccine_efficacy
        transmission_reduction <- rubella_params$transmission_reduction
        static_importation <- FALSE
        import_by_population <- FALSE
        days <- 365 * 25

        simulate_outbreak(R0 = R0,
            gamma = gamma,
            sigma = sigma,
            lambda_import = lambda_import,
            initial_immune = initial_immune,
            birth_rate = birth_rate,
            age_population = age_population,
            age_specific_mu_rate = age_specific_mu_rate,
            contact_matrix_load = contact_matrix_load,
            pathogen_x = "rubella",
            target_coverage = 0,
            vaccine_efficacy = vaccine_efficacy,
            transmission_reduction = transmission_reduction,
            static_importation = FALSE,
            import_by_population = import_by_population,
            days = 365 * 25) |>
            mutate(state = .x)
    })
holder <- bind_rows(holder)
us_df <- holder |> 
    mutate(state = "US") |>
    group_by(state, time) |> 
    summarize(across(population:female_repro_sus, sum),
              .groups = "drop")

pop_t0 <- us_df |> 
    filter(time == 1) |> 
    pull(population) |> 
    round() |> 
    prettyNum(big.mark = ",")

pop_tmax <- us_df |> 
    filter(time == max(time)) |> 
    pull(population) |> 
    round() |> 
    prettyNum(big.mark = ",")

total_births <- us_df |> 
    pull(births) |> 
    sum() |> 
    round() |> 
    prettyNum(big.mark = ",")

total_deaths <- us_df |> 
    pull(deaths) |> 
    sum() |> 
    round() |> 
    prettyNum(big.mark = ",")

mean_repro_women <- us_df |> 
    pull(female_repro_sus) |> 
    mean()
total_repro_women_pt <- round(mean_repro_women * 25) |> 
    prettyNum(big.mark = ",")
```


To support interpretation of the model results, we include some demographic details about the simulated US population over 25 years. At the start of the simulation, the US population was estimated at `r pop_t0` people and at the end of the simulation was estimated at `r pop_tmax` people. Over the 25-year simulation, we estimate `r total_births` births, which relate to the number of children eligible for vaccination. We estimate `r total_repro_women_pt` person-years of women of reproductive age (15-44 years) who were susceptible to rubella infection in the absence of routine childhood vaccination.  
